Higher-Order unification for free!

### 2024-04-13 paper/main.tex
sec 3.2 and 3.3 with spellchecker

\noindent
As we mentioned in section~\ref{sec:nutshell} the compiler
replaces terms in \maybeeta, \maybebeta and \notllambda with fresh
variables linked to the problematic terms. Terms in \maybebeta do not
need a link since \Ho{} variables faithfully represent
the problematic term thanks to their scope.

@@

\noindent
The right hand side of a link, the problematic term, can occur under binders.
To accommodate this situation the compiler wraps \elpiIn{baselink} using
the \elpiIn{inctx} container (see~\ref{data:inctx} also used for \elpiIn{subst}).

\begin{invariant}[Link left hand side]\label{inv:linklhs}
  The left hand side of a suspended link
  is a variable.
\end{invariant}

\noindent
New links are suspended by construction.
If the left hand side is assigned during a step, then 
the link is considered for progress and possibly eliminated.
This is discussed in \cref{sec:eta} and \cref{sec:beta}.

When detailing examples we write links as equations between two
terms under a context.
The equality sign is subscripted with
kind of \elpiIn{baselink}. For example $\linkbetaM{x}{A_x}{F_x~a}$ corresponds to:
\begin{elpicode}
abs x\ val (link-llam (uva A [x]) (app[uva F [x],con "a"]))
\end{elpicode}

@@

The simple compiler described in this section serves as a base for the
extensions in \cref{sec:llam,sec:eta,sec:beta}.
Its main task is to beta normalize the term and map one syntax tree to
the other.
In order to bring back the substitution from \Ho{} to \Fo{} the compiler
builds a ``memory map'' connecting the the kind of variables using routine
\ref{clause:malloc}.

The signature of the \elpiIn{comp} predicate below allows for the generation of
links (suspended unification problems) that play no role in this section
but play a major role in \cref{sec:llam,sec:eta,sec:beta}.
With respect to \cref{sec:problem-statement} the signature also allows
for updates to the substitution.

@@

\noindent
The code above uses that possibility
in order to allocate space for the variables, i.e. sets their memory
address to \elpiIn{none} (a details not worth mentioning in the
previous sections).
\todo{explain fold6}

@@

\noindent
In the code above the syntax \elpiIn{pi x y\..} is syntactic sugar for
iterated \elpiIn{pi} abstraction, as in \elpiIn{pi x\ pi y\..}.

The auxiliary function \elpiIn{close-links} tests if the bound variable
\elpiIn{v} really occurs in the link. If it is the case the link is wrapped into
an additional \elpiIn{abs} node binding \elpiIn{v}. In this way links generated
deep inside the compiled terms can be moved outside their original context
of binders.

>>>>

\noindent
As mentioned in section~\ref{sec:nutshell}, the compiler
replaces terms in \maybeeta, \maybebeta, and \notllambda with fresh
variables linked to the problematic terms. Terms in \maybebeta do not
need a link since \Ho{} variables faithfully represent
the problematic term thanks to their scope.

@@

\noindent
The right-hand side of a link, the problematic term, can occur under binders.
To accommodate this situation, the compiler wraps \elpiIn{baselink} using
the \elpiIn{inctx} container (see~\ref{data:inctx} also used for \elpiIn{subst}).

\begin{invariant}[Link left hand side]\label{inv:linklhs}
  The left-hand side of a suspended link
  is a variable.
\end{invariant}

\noindent
New links are suspended by construction.
If the left-hand side is assigned during a step, then 
the link is considered for progress and possibly eliminated.
This is discussed in \cref{sec:eta} and \cref{sec:beta}.

When detailing examples, we represent links as equations between two
terms under a context.
The equality sign is subscripted with the
kind of \elpiIn{baselink}. For example $\linkbetaM{x}{A_x}{F_x~a}$ corresponds to:
\begin{elpicode}
abs x\ val (link-llam (uva A [x]) (app[uva F [x],con "a"]))
\end{elpicode}

@@

The simple compiler described in this section serves as a base for the
extensions in \cref{sec:llam,sec:eta,sec:beta}.
Its main task is to beta normalize the term and map one syntax tree to
the other.
In order to bring back the substitution from \Ho{} to \Fo{} the compiler
builds a ``memory map'' connecting the kind of variables using routine
\ref{clause:malloc}.\noteD{non mi piace}

The signature of the \elpiIn{comp} predicate below allows for the generation of
links (suspended unification problems), which play no role in this section
but play a major role in \cref{sec:llam,sec:eta,sec:beta}.
With respect to \cref{sec:problem-statement}, the signature also allows
for updates to the substitution.

@@

\noindent
The code above uses that possibility
in order to allocate space for the variables, i.e. it sets their memory
address to \elpiIn{none} (a details not worth mentioning in the
previous sections).
\todo{explain fold6}

@@

\noindent
In the code above, the syntax \elpiIn{pi x y\..} is syntactic sugar for
iterated \elpiIn{pi} abstraction, as in \elpiIn{pi x\ pi y\..}.

The auxiliary function \elpiIn{close-links} tests if the bound variable
\elpiIn{v} really occurs in the link. If it does, the link is wrapped into
an additional \elpiIn{abs} node binding \elpiIn{v}. In this way links generated
deep inside the compiled terms can be moved outside their original context
of binders.
